\textsf{Are large language models superhuman chemists?}

### 2024-10-30 src/tex/appendix.tex
fix typos

\begin{itemize}
    \item \emph{End-to-end automation}. For model development, the evaluations must be run many times (e.g., on regular intervals of a training run).
    Approaches that rely on humans scoring the answers of a system\autocite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
    \item \emph{Careful validation by experts}. Manual curation is needed to minimize the number of incorrect or unanswerable questions.\autocite{northcutt2021pervasive}
    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\autocite{Frye_2023, Awg}
    \item \emph{Usable with models that support special treatment of molecules}. Some models, such as Galactica\autocite{taylor2022galactica}, use special tokenization or encoding procedures for molecules or equations.
    The benchmark system must encode the semantic meaning of various parts of the question or answer to support this.
    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or raw logits.
    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
    Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real-world chemistry, as well as higher-level university education, multiple-choice questions are seldom utilized.
    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond answering \gls{mcq}.
    \item \emph{Cover a diverse set of topics}. Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, and intuition.
    \item \emph{Cover a range of difficulty levels}. To allow for a continuous measure of improvement for a range of different (evolving) systems, a benchmark should cover a wide range of difficulty levels.
    \item \emph{Impossible to completely solve with current models}. A benchmark should contain questions that are impossible to solve with current models. If current models can solve all questions, the benchmark provides no useful signal.
\end{itemize}

@@

Reviewers also solved the questions to verify the answers. They also performed web searches to ensure questions were not easily found online. The reviewers often guided the revision process to ensure the question aligned with the guidelines. Questions that don't meet the criteria are either rejected or suggested for revision and most often, they are modified to a new question. Reviewers also provide feedback on the skill and difficulty annotations.

@@

The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.

@@

    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question. Note that the total count in this table is bigger than the \chembench corpus. This is because the same question can be annotated with two different \enquote{requires} skills, i.e. Reasoning and Calculation.} \\
    \label{tab:chembench_corpus_cognitive}

@@

            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\

@@

            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\

@@

            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\

@@

            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\

@@

\clearpage
\subsection{Model performance} \label{sec:model_performance_app}
We also evaluated the model performance on the entire \chembench corpus.
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/overall_performance.pdf}
    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered correctly by the models. Scores computed on the entire \chembench corpus.}
    \label{fig:barplot_all_correct_all_questions}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\subsection{Performance as a function of molecular features} \label{sec:molecular_features}
To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).
%The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals irregardless of molecular complexity). For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
\end{figure}

@@

\Cref{tab:refusal_counts_and_parsing} lists how many refusals were detected for the reponses of different models on \chembench. Overall, we find that refusals do not majorly affect the performance measured by \chembench.

@@

In our parsing workflow, we use pipelines based on regular expressions to extract the answers. In some cases, however, the answers are not directly extractable from the responses, for instance, when the model does not follow the formatting instructions. In these cases, we use a fallback mechanism to extract the answers. The fallback mechanism uses an \gls{llm} to extract the answers from the responses. The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer. We used \LlamaThreeSeventyBInstruct, accessed via the Groq API.
\Cref{tab:refusal_counts_and_parsing} tabulates the number of times the fallback mechanism was used for each model.

@@

\clearpage
\subsection{Implementation}
An overview of the benchmarking pipeline implemented in \chembench is shown in \Cref{fig:process}. More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
\begin{figure}
    \centering
    \includegraphics[height=.7\textheight]{figures/process.pdf}
    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}. \texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types. For each task type, appropriate prompts are generated and passed to the \gls{llm}. The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback. The parsed or extracted answers then undergo metric computation and refusal detection.}
    \label{fig:process}
\end{figure}

@@

\paragraph{Tool use}
In our study, humans were allowed to use tools for answering some questions.
They could also report what tools they used for answering questions. As \Cref{fig:tool_use} shows, the most commonly tool was some form of web search (which, according to the freetext responses, often was a multi-step process).

@@

\clearpage
\subsection{Tool augmented models} \label{sec:react-environment}
In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
We based the selection of these tools on the most used tools by humans (see \Cref{fig:tool_use}).
Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configuration.
We implemented the systems using Langchain\autocite{Chase_LangChain_2022} with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.

While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \gls{llm} without tools (\Cref{tab:performance_table}).
If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
However, the scores decreased for questions asking about the number of isomers or \gls{ghs} pictograms.
For the specific questions about converting \gls{iupac} names to \gls{smiles}, the results decreased notably despite the models having access to specific tools prepared for those questions.
By studying the reasoning path for these cases, we found that the error results from the models responding in a format that the LangChain framework with the default ReAct loop cannot handle.
This indicates that agent frameworks need optimization to be more robust. It involves not only equipping the \glspl{llm} with tools but also necessitates engineering efforts to create robust systems.

>>>>

\begin{itemize}
    \item \emph{End-to-end automation}. For model development, the evaluations must be run many times (e.g., on regular intervals of a training run).
    Approaches that rely on humans scoring the answers of a system\autocite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
    \item \emph{Careful validation by experts}. Manual curation is needed to minimize the number of incorrect or unanswerable questions.\autocite{northcutt2021pervasive}
    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\autocite{Frye_2023, Awg}
    \item \emph{Usable with models that support special treatment of molecules}. Some models, such as Galactica\autocite{taylor2022galactica}, use special tokenization or encoding procedures for molecules or equations.
    The benchmark system must encode the semantic meaning of various parts of the question or answer to support this.
    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or raw logits.
    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
    Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real-world chemistry, as well as higher-level university education, multiple-choice questions are seldom utilized.
    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond answering \gls{mcq}.
    \item \emph{Cover a diverse set of topics}. Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities}, the topics covered by a chemistry benchmark must be very diverse.
    \item \emph{Cover diverse skills}. To holistically judge performance, it is important to cover questions that rely on reasoning, calculation, knowledge, and intuition.
    \item \emph{Cover a range of difficulty levels}. To allow for a continuous measure of improvement for a range of different (evolving) systems, a benchmark should cover a wide range of difficulty levels.
    \item \emph{Impossible to completely solve with current models}. A benchmark should contain questions that are impossible to solve with current models. The benchmark provides no useful signal if current models can solve all questions.
\end{itemize}

@@

Reviewers also solved the questions to verify the answers. They also performed web searches to ensure questions were not easily found online. The reviewers often guided the revision process to ensure the question aligned with the guidelines. Questions that don't meet the criteria are either rejected or suggested for revision, and most often, they are modified to a new question. Reviewers also provide feedback on the skill and difficulty annotations.

@@

The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong.

@@

    \caption{\textbf{Examples for each required skill considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question. Note that the total count in this table is bigger than the \chembench corpus. This is because the same question can be annotated with two different skills, e.g., Reasoning and Calculation.}    \label{tab:chembench_corpus_cognitive} \\

@@

            \multicolumn{1}{c}{\textbf{Knowledge} \variable{output/knowledge_count.txt} Questions} \\

@@

            \multicolumn{1}{c}{\textbf{Reasoning} \variable{output/reasoning_count.txt} Questions} \\

@@

            \multicolumn{1}{c}{\textbf{Intuition} \variable{output/intuition_count.txt} Questions} \\

@@

            \multicolumn{1}{c}{\textbf{Calculation} \variable{output/calculation_count.txt} Questions} \\

@@

\clearpage
\subsection{Model performance} \label{sec:model_performance_app}
We also evaluated the model performance on the entire \chembench corpus.
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions the models answered correctly.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/overall_performance.pdf}
    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions the models answered correctly. Scores computed on the entire \chembench corpus.}
    \label{fig:barplot_all_correct_all_questions}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\subsection{Performance as a function of molecular features} \label{sec:molecular_features}
To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals regardless of molecular complexity). For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
\end{figure}

@@

\Cref{tab:refusal_counts_and_parsing} lists how many refusals were detected for the responses of different models on \chembench. Overall, we find that refusals do not majorly affect the performance measured by \chembench.

@@

Our parsing workflow uses pipelines based on regular expressions to extract the answers. In some cases, however, the answers are not directly extractable from the responses, for instance, when the model does not follow the formatting instructions. In these cases, we use a fallback mechanism to extract the answers. The fallback mechanism uses an \gls{llm} to extract the answers from the responses. The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer. We used \LlamaThreeSeventyBInstruct, accessed via the Groq API.
\Cref{tab:refusal_counts_and_parsing} tabulates the number of times the fallback mechanism was used for each model.

@@

\clearpage
\subsection{Implementation}
An overview of the benchmarking pipeline implemented in \chembench is shown in \Cref{fig:process}. More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
\begin{figure}
    \centering
    \includegraphics[height=.7\textheight]{figures/process.pdf}
    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}. \texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types. For each task type, appropriate prompts are generated and passed to the \gls{llm}. The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
    The regex-based parsing is elaborate and can handle special cases such as scientific notation or Roman numerals.
    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback. The parsed or extracted answers then undergo metric computation and refusal detection.}
    \label{fig:process}
\end{figure}

@@

\paragraph{Tool use}
In our study, humans were allowed to use tools for answering some questions.
They could also report what tools they used for answering questions. As \Cref{fig:tool_use} shows, the most common tool was some form of web search (which, according to the free text responses, often was a multi-step process).

@@

\clearpage
\subsection{Tool augmented models} \label{sec:react-environment}
In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
For this, we investigated the two models that showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
We based the selection of these tools on the most used tools by humans (see \Cref{fig:tool_use}).
Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
These conversion tools allow us to understand better how the agents perform for specific questions in this agent environment configuration.
We implemented the systems using Langchain\autocite{Chase_LangChain_2022} with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.

While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \gls{llm} without tools (\Cref{tab:performance_table}).
If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
However, the scores decreased for questions about the number of isomers or \gls{ghs} pictograms.
For the specific questions about converting \gls{iupac} names to \gls{smiles}, the results decreased notably despite the models having access to specific tools prepared for those questions.
By studying the reasoning path for these cases, we found that the error results from the models responding in a format that the LangChain framework with the default ReAct loop cannot handle.
This indicates that agent frameworks need optimization to be more robust. It involves not only equipping the \glspl{llm} with tools but also necessitates engineering efforts to create robust systems.
